---
title: "The Meta-Analysis of Nonpoint Source BMPs: Can We Improve the Empirical Knowledge Associated with BMP Water Quality Improvement?"
author: "Michael Schramm, Duncan Kikoyo, Janelle Wright, Shubham Jain"
institute: "Texas Water Resources Institute, Texas A&M AgriLife Research.\\hfill \\break Project funding provided by the Texas State Soil and Water Conservation Board. \\hfill \\break 2024 AWRA, UCOWR, & NIWR Joint Water Resources Conference"
date: 8/30/2024
format: 
  beamer:
    theme: moloch
    themeoptions: progressbar=frametitle
    aspectratio: 169
    keep-tex: true
    include-in-header: preamble.tex
    bibliography: bibliography.bib
    cite-method: biblatex
    biblio-style: apa
    include-before-body: before-body.tex
    template-partials: 
      - biblio.tex
    

---

```{r setup}
library(targets)
library(ggtext)
library(kableExtra)
library(patchwork)
library(ragg)
library(targets)
library(tidyverse)

## flowchart
library(ggdag)
library(rsvg)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "ragg_png",
                      echo = FALSE,
                      ## figure chunk options
                      fig.width = 6,
                      dpi = 300)

## make fonts available
font_hoist <- function(family, silent = FALSE) {
  font_specs <- systemfonts::system_fonts() |> 
    filter(family == {{family}}) |> 
    mutate(family = paste(family, style)) |> 
    select(plain = path, name = family)
  
  purrr::pwalk(as.list(font_specs), systemfonts::register_font)

  if (!silent)  message(paste0("Hoisted ", nrow(font_specs), " variants:\n",
                               paste(font_specs$name, collapse = "\n")))
}

font_hoist("Manrope")

## grab output from targets to use in manuscript
store <- "../_targets"
```


## Introduction

### Purpose

 - Assess scope and scale of published water quality BMP field studies.
 - Assess mean reductions in pollutant *concentrations* in BMP studies.
 - Methods and results fully detailed in Frontiers article \footfullcite{schrammMetaanalysisImpactsBest2024}

## Methods

### Systematic Review
 - Followed systematic review guidance in Collaboration for Environmental Evidence \footfullcite{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}
  - Review identified **90 studies** and **642 effect sizes**
 - Collected data included: *Publication date, measured parameters, runoff source, BMP type, study scale, location, study length, measurement units, control and treatment sample size, standard error, standard deviation, mean concentration, etc.*
 
 - Raw data available on Zenodo \footfullcite{kikoyoTWRINonpointSource2023} 

## Methods

```{r}
#| id: bmpstruct
#| fig-width: 6
#| fig-height: 3
dag <- dagify(
  manure_management ~ livestock,
  prescribed_grazing ~ livestock,
  livestock_exclusion ~ livestock,
  alternate_water ~ livestock,
  reduced_till ~ crop,
  no_till ~ crop,
  nutrient_management ~ crop,
  cover_crop ~ crop,
  crop ~ management,
  livestock ~ management,
  sand_filter ~ filtration,
  swale ~ filtration,
  filter_strip ~ filtration,
  infil_trench ~ infiltration,
  infil_chamber ~ infiltration,
  infil_basin ~ infiltration,
  retention_basin ~ detention,
  sediment_trap ~ detention,
  detention_basin ~ detention,
  tree_filter ~ treatment,
  bioreactor ~ treatment,
  wetland ~ treatment,
  bioret ~ treatment,
  dtr ~ drainage,
  proprietary ~ drainage,
  green_roof ~ drainage,
  gss ~ drainage,
  controlled_drainage ~ drainage,
  filtration ~ structural,
  infiltration ~ structural,
  detention ~ structural,
  treatment ~ structural,
  drainage ~ structural,
  labels = c(
    "manure_management" = "Manure Management",
    "prescribed_grazing" = "Prescribed Grazing",
    "livestock_exclusion" = "Livestock Exclusion",
    "alternate_water" = "Alternate Water",
    "reduced_till" = "Reduced Till",
    "no_till" = "No Till",
    "nutrient_management" = "Nutrient Management",
    "cover_crop" = "Cover Crop",
    "crop" = "Crop",
    "livestock" = "Livestock",
    "sand_filter" = "Sand Filter",
    "swale" = "Swale",
    "filter_strip" = "Filter Strip",
    "infil_trench" = "Infiltration Trench",
    "infil_chamber" = "Infiltration Chamber",
    "infil_basin" = "Infiltration Basin",
    "retention_basin" = "Retention Basin",
    "sediment_trap" = "Sediment Trap",
    "detention_basin" = "Detention Basin",
    "tree_filter" = "Tree Filter",
    "bioret" = "Bioretention",
    "bioreactor" = "Bioreactor",
    "wetland" = "Wetland",
    "dtr" = "Drainage Tailwater Recovery",
    "proprietary" = "Proprietary",
    "green_roof" = "Green Roof",
    "gss" = "Grade Stabilization Structure",
    "controlled_drainage" = "Controlled Drainage",
    "filtration" = "Filtration",
    "infiltration" = "Infiltration",
    "detention" = "Detention",
    "treatment" = "Treatment",
    "drainage" = "Drainage",
    "management" = "Management\nBMPs",
    "structural" = "Structural\nBMPs"
  )
) |> 
  tidy_dagitty(layout = "tree", circular = FALSE) |>
  mutate(rotate = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ 90,
    .default = 0
  ),
  node_size = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ NA,
    .default = 20
  ),
  hjust = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ 0,
    .default = 0.5
  ),
  nudge_y = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ 1,
    .default = 0
  )) 
  
ggplot(dag, aes(x = -y, y = x, xend = -yend, yend = xend)) +
  geom_dag_node(aes(size = node_size), color = "wheat") +
  geom_dag_edges_link() +
  geom_dag_text(data = dag |> 
                  filter(nudge_y == 0),
                aes(angle = rotate, hjust = hjust, label = label), size = 2, color = "black") +
  geom_dag_text(data = dag |>
                  filter(nudge_y == 1),
                aes(#angle = rotate, 
                    hjust = hjust, label = label), size = 2, color = "black",
                nudge_x = -0.1) +
  scale_size_area(max_size = 20) +
  scale_y_continuous(limits = c(-11,17),
                     expand = expansion(0,0)) +
  scale_x_continuous(limits = c(-2.5, 1),
                     expand = expansion(0,0)) +
  theme_dag() +
  theme(legend.position = "none")

```

## Methods

### Meta-Analysis

 - Describe overall effect and heterogeneity
 - Weighted multilevel random effects models
    - Weights individual effect sizes based on sampling variance
 - Model terms:
    - Influent pollutant concentration
    - BMP category
    - Aridity index: $\frac{\bar{P}}{\bar{PET}}$ \footfullcite{zomerVersionGlobalAridity2022}
    - Fixed effect interactions
    - Nested random effects $(Effect|Study)$

## Methods

### Meta-Analysis

 - Dependent variable: log ratio of means
 
$$
ROM_i = ln(\frac{x_{i,control}}{x_{i,trmt}}) = ln(x_{i,control}) - ln(x_{i,trmt})
$$
 - Separate models for each parameter of interest: **fecal indicator bacteria** (FIB), **total nitrogen** (TN), **total phosphorus** (TP), **total suspended sediment** (TSS), **orthophosphate** (PO~4~), **dissolved inorganic nitrogen** (DIN).

## Study Scale & Source

```{r}
#| label: p1
#| fig-height: 3
#| fig-width: 8
#| out-width: "100%"
#| fig-dpi: 300

tar_read(bmp_summary, store = store) &
  theme(legend.text = element_text(size = rel(0.65)),
        legend.title = element_text(size = rel(0.65)),
        legend.position = c(0.8, 0.65),)
```

- Most studies were lot/field scale and investigated urbanized runoff.
- FIB studies tended toward larger catchments.

## Publication Date and Study Length



```{r}
#| label: p2
#| fig-dpi: 300
#| fig-width: 6
#| fig-height: 6
#| out-width: "50%"
#| 
tar_read(bmp_temporal, store = store)$p1
```


## Spatial Distribution

```{r}
#| label: p3
#| fig-dpi: 300
#| fig-width: 6.5
#| fig-height: 6.5
#| out-width: "50%"
#| 
tar_read(study_map, store = store)
```


## Results


::::: {.columns}

:::: {.column width=60%}

```{r}
#| label: p4
#| fig-height: 5.75
#| fig-width: 6.5
#| out-width: "100%"
#| fig-dpi: 300
tar_read(overall_effect, store = store)
```
::::


:::: {.column width=40%}

### Overall Effect

* Strong evidence across studies for mean reductions in:
  + **FIB (57%)** 
  + **TN (34%)** 
  + **TP (32%)** 
  + **TSS (81%)**

* Prediction intervals provide low certainty about the performance of future *individual* studies.


::::

:::::


## Attempts to Explain Variation in BMP Performance

::::: {.columns}

:::: {.column width=40%}

```{r}
tibble::tibble("Parameter" = c("FIB", "TP", "PO\\textsubscript{4}"),
               "Moderators" = kableExtra::linebreak(c("Influent Concentration,\nAridity × BMP Subcategory",
                                "Influent Concentration",
                                "Influent Concentration"))) |> 
  kableExtra::kbl(escape = FALSE)
```

Influent concentration + Aridity×BMP Subcategory explained ~89% of effect size variance in FIB models.

::::

:::: {.column width=60%}

```{r}
#| label: p5
#| fig-width: 6
#| fig-height: 4
#| out-width: "100%"
#| fig-dpi: 300
tar_read(bac_marginal_means, store = store) &
  theme(legend.position = "none")
```

::::

:::::


## Attempts to Explain Variation in BMP Performance

::::: {.columns}

:::: {.column width=50%}

```{r}
tibble::tibble("Parameter" = c("FIB", "TP", "PO\\textsubscript{4}"),
               "Moderators" = kableExtra::linebreak(c("Influent Concentration,\nAridity × BMP Subcategory",
                                "Influent Concentration",
                                "Influent Concentration"))) |> 
  kableExtra::kbl(escape = FALSE)
```

Influent concentration explained 12% and 35% of effect size variance in TP and PO~4~ models.

::::

:::: {.column width=50%}

```{r}
#| label: p6
#| fig-width: 6
#| fig-height: 4
#| out-width: "100%"
#| fig-dpi: 300
tar_read(phos_marginal_means, store = store) &
  theme(legend.position = "none")
```

::::

:::::

## Overview

* Few long-term studies (>4 year).
* Studies clustered in the mid-Atlantic, low spatial variability.
* Studies demonstrated mean reductions in FIB, TN, TP, and TSS.
* Did not find evidence for mean reductions in DIN, PO~4~.
* Influent concentration has strong influence on FIB, TP, and PO~4~ performance.
* Failed to explain majority of variance in nutrient and TSS models. **(Why???)**

## Lessons Learned

**Issues in Published BMP Studies**

* Very inconsistent reporting of non-manipulated study conditions.
* Earlier studies tended to report BMP efficiency only.
* Many studies fail to report sample size, variance, or uncertainty measures (59% of identified effect sizes).
* There is an opportunity to develop reporting protocols for BMP studies. See Eagle et al. \footfullcite{eagleMetaAnalysisConstrained2017}
  - Environmental variables, reported metrics, regression model reporting...


## Lessons Learned

**Study/Systematic Review Issues**

* Broad study design (focus on specific BMP types).
* Keyword comprehensiveness.
* *metafor* R package facilitates meta-analysis.\footfullcite{viechtbauerConductingMetaanalysesMetafor2010} 
* Developing and training students/staff on data extraction is difficult.
* Data entry interfaces that reinforce data consistency.
    

## Thank You

**Contact:** michael.schramm@ag.tamu.edu

**Acknowledgements:** Texas State Soil and Water Conservation Board for funding.

**Paper:** Schramm M, Kikoyo D, Wright J and Jain S (2024) A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration. Front. Water 6:1397615. doi: 10.3389/frwa.2024.1397615

**Data:** Kikoyo D, Jain S, Wright J, and Schramm M. (2024). TWRI NPS-BMP-DB (v1.05). Zenodo. doi: 10.5281/zenodo.10451303



