---
output: github_document
---

# A  meta-analysis  of  the  impacts  of  best management  practices  on  nonpoint  source pollutant  concentration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This repository contains the data and analysis used in the draft manuscript *"A  meta-analysis  of  the  impacts  of  best management  practices  on  nonpoint  source pollutant  concentration,"* by Schramm et al.

This project uses `{targets}` as the main organizing workflow. The `_targets.R` file defines the steps of the project workflow amd functions for those setps are stored in the `R/` directory. `renv` is used to manage package versions. To reproduce locally, clone or download this project to your computer. Open the project with RStudio or other IDE.

In the console:

```r
renv::restore()
```

then use `targets::tar_make()` to execute the project pipeline. Each step in the pipeline can be accessed with the `targets::tar_read()` function. For example, `targets::tar_read(tn_model)` will access the model object we fit to the total nitrogen data.

## Funding Statement

This project was funded by a nonpoint source grant from the Texas State Soil and Water Conservation Board.

## Working Environment Info

```{r}
renv::diagnostics()
```